P4IRS: An Intermediate Representation and Compiler for Parallel and Performance-Portable Particle Simulations

### 2023-11-19 manuscript.tex
Contact history text
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

>>>>

When implementing DEM kernels such as the Linear Spring-Dashpot, properties computed during two contacted particles must be used across subsequent time-step interactions whenever they happen.
For this reason, P4IRS supports contact properties, which are properties tighted to a contact in the simulation.
These can be simply added through the \emph{add\_contact\_property} method as can be seen in \autoref{lst:contactprops1} for the sticking, tangential spring displacement and the impact velocity magnitude:

\begin{lstlisting}[language=Python,
		   label={lst:contactprops1},
		   caption={Setup example for contact properties.}]
psim.add_contact_property('is_sticking', pairs.int32(), 0)
psim.add_contact_property('tangential_spring_displacement', pairs.vector(), [0.0, 0.0, 0.0])
psim.add_contact_property('impact_velocity_magnitude', pairs.real(), 0.0)
\end{lstlisting}

These properties can be used directly in kernels in the same way as feature properties.
\autoref{lst:contactprops2} shows an example of their usage in the Linear Spring-Dashpot kernel.
Note that when such properties are not stored for the particle pair, then the default values are always used.

\begin{lstlisting}[language=Python,
		   label={lst:contactprops2},
		   caption={Setup example for contact properties.}]
def linear_spring_dashpot(i, j):
    # ...
    tan_spring_disp = tangential_spring_displacement[i, j]
    impact_vel_magnitude = impact_velocity_magnitude[i, j]
    impact_magnitude = select(impact_vel_magnitude > 0.0, impact_vel_magnitude, length(rel_vel))
    # ...
    tangential_spring_displacement[i, j] = new_tan_spring_disp
    impact_velocity_magnitude[i, j] = impact_magnitude
    is_sticking[i, j] = n_sticking
\end{lstlisting}

To provide fast and efficient accesses of contact properties when using neighbor-lists, P4IRS keep the particles in the same indexes in both data structures.
This prevents a lookup operation within the kernel for finding the contact index, and also keeps data access contiguous for the contact history data structure.
When Linked Cells are used without a Verlet Lists, such optimizations are not possible, which should significantly impact the performance for the kernel.
